---
title: "q3"
author: "Ziyi Wang"
date: "2023-04-30"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyverse)
```

#### 3. Compute posterior summaries and 95% credible intervals of $\gamma$, the fixed effects associated with the covariates in the model. Using the estimated Bayesian model, answer the following questions: (1) are there seasonal differences in hurricane wind speeds, and (2) is there evidence to support the claim that hurricane wind speeds have been increasing over the years?

# TASK

## Compute posterior summaries of $\gamma$

```{r}
# load data from Q2 gibbs
dataRDS = readRDS("./data/new_gibb5000.rds")

#head(dataRDS) #dont do this, crazy

names(dataRDS)
# beta = dataRDS$B
# mu = dataRDS$mu
# sigmasq = dataRDS$sigmasq
# SIGMA = dataRDS$SIGMA
gamma = dataRDS$gamma

# every element in the gamma list i.e.[[i]] is the ith gibbs run
gamma[[1]] 
as.matrix(gamma[[2]])

L = length(gamma)

# get last n gamma
n = 1000
df_gama = NULL

for (i in 1:n) {
    new = gamma[[L + 1 - i]]
    df_gama <- rbind(df_gama, c(L + 1 - i, t(as.matrix(new))))
}
df_gama2 = as.data.frame(df_gama)
names(df_gama2) = c("id",names(gamma[[L]]))
head(df_gama2)

```

## plot month

```{r}
df_month <- df_gama2 %>%
  select(-season, -TS, -ET, -SS, -NR) %>%
  gather(key = "month", value = "value", -id)
head(df_month)

ggplot(df_month, aes(x = id, y = value)) + 
  geom_line(aes(color = month, linetype = month)) 
#+ scale_color_manual(values = c("darkred", "steelblue", "red"))
```

## plot year

```{r}
df_year <- df_gama2 %>%
  select(id, season) %>%
  gather(key = "name", value = "year", -id)
head(df_year)

ggplot(data = df_year, aes(x = id, y = year)) +
  geom_line()
```

## plot type

```{r}
#theme_set(theme_minimal())
df_type <- df_gama2 %>%
  select(id, TS, ET, SS, NR) %>%
  gather(key = "type", value = "value", -id)
head(df_type)

ggplot(df_type, aes(x = id, y = value)) + 
  geom_line(aes(color = type, linetype = type)) 
#+ scale_color_manual(values = c("darkred", "steelblue", "red"))
```

## Compute 95% credible intervals of $\gamma$, the fixed effects associated with the covariates in the model.

```{r}
cred_int <- apply(df_gama2, 2, quantile, probs = c(0.025, 0.975))
cred_int
```

## (1) are there seasonal differences in hurricane wind speeds

```{r}
# E.X. May - April
df_gama2$May_April <- (df_gama2$May - df_gama2$April)
quantile(df_gama2$May_April, probs = c(0.025, 0.975))
# include 0, meaning...no diff betw. May & April
```

```{r}
# E.X. June - May
df_gama2$J_M <- (df_gama2$June - df_gama2$May)
quantile(df_gama2$J_M, probs = c(0.025, 0.975))
# include 0, meaning...no diff betw. June & May
```

```{r}
# E.X. Aug - July
df_gama2$A_J <- (df_gama2$August - df_gama2$July)
quantile(df_gama2$A_J, probs = c(0.025, 0.975))
# include 0, meaning...no diff betw. Aug & July
```

## summer month
```{r}
df_summer =  df_gama2 %>% select(June, July, August)
df_nosummer =  df_gama2 %>% select(April, May, September,
                                   October, November, December)
     
summer_wd = rowSums(df_summer)/3
nosummer_wd = rowSums(df_nosummer)/6

head(summer_wd - nosummer_wd)
quantile(summer_wd - nosummer_wd, probs = c(0.025, 0.975))
```

## summer season: 7,8,9,10
```{r}
df_summer =  df_gama2 %>% select(July, August, September, October)
df_nosummer =  df_gama2 %>% select(April, May, June, November, December)
     
head(df_summer)

summer_wd = rowSums(df_summer)/4
nosummer_wd = rowSums(df_nosummer)/5

head(summer_wd - nosummer_wd)
quantile(summer_wd - nosummer_wd, probs = c(0.025, 0.975))

```


## (2) is there evidence to support the claim that hurricane wind speeds have been increasing over the years

H0: gamma14 \<= 0; H1: gamma14 \> 0

```{r}
quantile(df_year$year, probs = c(0.025, 0.975))
#         2.5%        97.5% 
# -0.003099822 -0.001501638
```

The gamma associated with type (Nature of the hurricane) is smaller than 0, fail to reject H1. No evidence to support the claim that hurricane wind speeds have been increasing over the years.
